### 2020-12-22 analysis/paper_calenviro.Rmd
descriptive statistics, paper progress

---
title: "paper_calenviro"
author: "David Purucker"
date: "12/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

``` {regression table}
library(stargazer)
model1 <- lm(lbw~ozone+pm2.5+dieselpm+drinkingwater+pesticides+toxrelease+traffic+
               cleanupsite+groundwater+hazwastes+impwater+solidwaste, data=ces3_complete)
model2 <- lm(lbw~ozone+pm2.5+dieselpm+drinkingwater+pesticides+toxrelease+traffic+
               cleanupsite+groundwater+hazwastes+impwater+solidwaste+poverty, data=ces3_complete)
model3 <- lm(lbw~ozone+pm2.5+dieselpm+drinkingwater+pesticides+toxrelease+traffic+
               cleanupsite+groundwater+hazwastes+impwater+solidwaste+pctnonwhite, data=ces3_complete)
model4 <- lm(lbw~ozone+pm2.5+dieselpm+drinkingwater+pesticides+toxrelease+traffic+
               cleanupsite+groundwater+hazwastes+impwater+solidwaste+poverty+pctnonwhite, data=ces3_complete)
screenreg(lapply(list(model1, model2, model3, model4), convertModel),
          caption="OLS regression models predicting low birth-weight",
          caption.above = TRUE)
stargazer(model1)

stargazer(model.svy1, model.svy2, model.svy3, model.svy4, model.svy5, type="html", digits = 3, covariate.labels = c("GPA", "Black/African American", "Latino", "Asian/Pacific Islander", "Other", "American Indian/Native American", "Male", "Grade in school", "Honor Society", "Parent Income", "Uses Alcohol", "Smokes", "Band/Choir", "Academic club", "Sports"), notes.align = "l", notes.append = TRUE, keep.stat = c("n","rsq"), notes = c("Standard errors in parenthesis.", "The reference category for race is white and for gender is female.", "GPA is based on self reported grades and is out of a 4.0 scale.", "Sports measures the number of sports a student is involved in from 0-6.", "Alcohol use, smokes, and other extracurricular activites are true/false variables."))

stargazer(model1, model2, model3, model4, model5, type="html", digits = 3)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.