### 2021-01-04 resources/paper_calenviro.Rmd
finished paper, except for formatting (how do I get it to output a PDF?)

---
title: "OLS Regression Project"
author: "David Purucker"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, message=FALSE, include=FALSE)
load("~/Desktop/sociology/soc_year2/Soc 412:512 (Stats II)/calenviro_proj/output/ces3_complete.RData")
if(!require(stargazer)) {
  install.packages("stargazer", "texreg")
}
```

# Introduction

What is the relationship between hazarodus environmental pollutants, socioeconomic factors, and low birth-weight? Birth-weight is an important indicator of overall population health. Babies weighing less than 2,500 grams (~5.5 lbs.) at birth are classified as low birth-weight (LBW). LBW is a useful proxy for studying pre-term birth (births occurring before 37 weeks of gestation), which is a leading cause of worldwide infant and child mortality (WHO 2018). LBW babies are also at higher risk of suffering health problems later in life, such as asthma (Nepomnyaschy and Reichman 2006), coronary heart disease and type 2 diabetes (Barker et al. 2002), and developmental delays (Bhutta et al 2002). LBW risk is increased by social and environmental stressors, including poverty (Martinson and Reichman 2015), exposure to toxic air contaminants (Ghosh et al. 2012), pesticides (Harley et al, 2011), and living in close proximity to freeways (Laurent et al 2013). Chronic health conditions associated with LBW may in turn make LBW infants more sensitive to environmental pollution after birth. LBW is more prevalent among African-American women than white and non-white Hispanic women, and is a contributing factor to the disproprotionate rates of maternal and infant mortality among Black women in the United States (Lu and Halfon 2003). LBW in the United States is concentrated in communities of color (Sims, Sims, and Bruce 2008). Identifying pollutant predictors for LBW is therefore relevant for policymaking aimed at achieving environmental justice.

To study this question, I used CalEnviroScreen 3.0, a public health modeling tool designed to assess the 'cumulative impacts' of environmental and social stressors on Californians (Rodriquez and Zeise 2017). Cumulative impact refers to the totality of health risks affecting a population, including both environmental and socio-economic conditions. In this analysis, OLS regrssion models are used to analyze the correlations of various pollution measures from CalEnviroScreen with the incidence of LBW at the census tract level, to determine the prevalence of environmental and social correlates of LBW in the CES data, and understand the strengh of different LBW effects of each stressor. Findings indicate a positive and moderately strong independent effect of ozone air pollution on LBW, and similarly positive, though weaker, effects from particulate matter air pollution. These findings suggest that policies aimed at curtailing air pollution, especially ozone, will have the most beneficial effect on reducing LBW and its associated public health effects.

# Literature review

The relationship between air pollution and low birth-weight has been very well-studied. In a recent systematic meta-analysis, Bekkar et al. (2020) evaluate the results of 57 studies conducted between 2007 and 2019 in the United States that consider the relationship of adverse pregnancy outcomes (low birth-weight, preterm birth, and stillbirth) and environmental conditions associated with climate change, including heat exposure and the two forms of air pollution, ozone and PM2.5. Their meta-analysis, which includes data on nearly 38.8 million births, indicate a strong positive relationship between heat and both pollutants with adverse birth outcomes. 25 out of 29 studies in the review found a positive association between PM2.5 or ozone exposure with LBW. These studies also indicate that some subpopulations are at particular risk of pollution-associated LBW, including asthmatics and Black women. The large sample size and geographic range of this review, encompassing every region of the United States, strongly supports the prediction that a positive relationship between LBW and air pollution variables will be found in the CalEnviroScreen 3.0 dataset. [note Bekkar et al.'s scale of analysis—-do most of their sampled studies address LBW at the county level, cities, tracts, ???]

The connection between LBW and other forms of pollution included in the CalEnviroScreen data is also well-established. LBW correlates have been found in studies considering proximity to freeway traffic (Laurent et al. 2013, Ponce et al. 2005), pesticides (Harley et al. 2011), diesel particulate matter (Laurent et al. 2016), contaminated drinking water (Currie et al. 2013) and proximity to hazardous waste sites (Berry and Bove 1997). Finally, social determinants of health such as porverty and race also have a well-established relationship with LBW (Lu and Hafton 2003, Martinson and Recihman 2016).

# Data and methods

CalEnviroScreen (https://oehha.ca.gov/calenviroscreen/report/calenviroscreen-30) , or CES, is a science-based screening tool developed by the California EPA to assess the cumulative burden of environmental pollutants and social stressors on populations around the state. CES compiles data on variables related to cumulative burden for 8,035 census tracts in California. This data is amalgamated from a variety of local, state, and federal government sources, depending on the environmental or social indicator in question (Rodriquez and Zeise 2017). Data for environmental burdens is interacted with population characteristics like poverty and cardiovascular disease to calculate a cumulative burden score for each census tract. Given California's size and its economic and geographic diversity, it seems reasonable to assume that trends in CalEnviroScreen data can be generalized to predict public health outcomes in other settings, at least within developed countries. Epidemiological researchers and policymakers are increasingly interested in statistical techniques that quantify such cumulative impacts, and CalEnviroScreen has been rated positively for this purpose (Greenfield, Rajan, and McKone 2017).

Data on environmental burdens is provided by various boards and departments within the California EPA, as well as the California Department of Transportation and the federal EPA. Data on population characteristics come from the US Census Bureau and the California Department of Public Health. Some data was collected at the tract level, while other data was originally collected at different scales and then adjusted to fit tracts. For this analysis, all individual indicators of environmental hazard are retained as independent variables. Population characteristics included in this analysis are low birth-weight (the dependent variable), poverty, and race/ethnicity. The total cumulative burden score (CES score) is not retained as a variable in this analysis, because it combines together environmental and social factors that I wish to disaggregate and test for their associations with a specific public health condition, low birth-weight. Including the CES score in regression models would also pose a risk of double-counting. Some work was conducted to clean this data for the analysis. Missing values were removed for PM2.5 pollution, traffic, drinking water, poverty, 
pct. nonwhite, and LBW variables, meaning that any 0-unit scores in this analysis are 'true' zeroes, and not placeholders for missing, unrecorded data. After removing missing data, the final sample for this analysis is 7,750 tracts.

The dependent variable of this analysis, low birth-weight, is measured as the percentage of live, singleton births with LBW (that is, weighing less than 5.5 lbs.) in the tract each year, averaged over 2006-12. The data excludes tracts that had fewer than 50 live births during the seven years in which the data was collected. Poverty is included as a control variable. Poverty is measured as the estimated percent of the population living below two times the federal poverty level, from 2011-15 (American Community Survey data). Because of its strong known associations with LBW and frequent coincidence with environmental stressors, leaving poverty out of my models could obscure significant independent effects of particular pollution exposures. Finally, a simple measure of tract-level race/ethnicity is included because of the implications of this analysis for environmental and racial justice. This variable is constructed as the percent of the tract population that is not white, exclusive of Hispanics.

Indicators classified as pollution 'exposures' in CES includes ozone, measured as the amount of daily maximum 8-hour ozone concentration between May-October (parts per million), averaged over 2012-14; PM2.5 (fine particulate matter, ~2.5 microns), measured as an annual mean concentration (μg/m3) between 2012-14; particulate matter from on- and off-road diesel exhaust, measured as a spatial distribution of diesel PM emissions (kg/day) on a July day in 2012; an index of drinking water contaminants (including nitrate, perchlorate, and animal waste), measured between 2005-13; pesticides, measured as total pounds of selected active pesticide ingredients used in production-agriculture per square mile in the census tract, averaged over 2012-14; toxic releases, measured as toxicity-weighted concentrations of modeled chemical releases to air from facility emissions and off-site incineration, averaged over 2011-13; and traffic density, measured as vehicle-kilometers per hour per road length, within 150 meters of the census tract boundary in 2013. Indicators classified within the category of 'environmental effects' include brownfield clean-up sites, measured as the sum of sites per tract in 2016, weighted for proximity to populated census blocs; groundwater threats (such as leaking storage tanks) measured as the sum of sites per tract in 2016, weighted for cleanup status and proximity to population; hazardous wastes, measured as the sum of hazardous waste facilities per tract between 2012-2016; impaired water bodies, measured as the sum of water bodies designated as polluted in 2012; and solid waste sites and facilities, measured as the sum of such sites per tract in 2016.

This analysis uses an OLS regression modeling approach to analyze the relative effects of different environmental and social stressors on the incidence of LBW in the CES data. After summarizing the distribution of the dependent and key independent variables in the data, I construct four regression effects models. First, I construct a model that includes all of the environmental effects, without controlling for the effects of race and poverty. Next, I construct two models that control for the effect of poverty and race separately. Finally, I combine both race and poverty into a model alongisde all of the environmental effects variables, to determine the relative association of each.

# Descriptive statistics

The histogram in figure 1 shows the distribution of the dependent variable: low birth-weight rates across census tracts. LBW displays a symmetrical, normal distribution in this data. Most of the cases (tracts) fall between 4 and 7 per cent of total births. The graph should be interpreted as follows: there are about 500 census tracts with an LBW rate between 2 and 3 per cent of total births, about 1250 tracts with an LBW rate between 3 and 4 percent of total births, about 2000 tracts with an LBW rate between 4 and 5 per cent of total births, and so on. The mean of the distribution is 4.979 (per-cent births LBW), while the median is 4.92, meaning the distribution is very slightly right-skewed. There are a very small number of outlier tracts with extremely high rates of LBW (the maximum tract value is 14.89), as well as some tracts with a zero per-cent rate of LBW.

The scatterplot in figure 2 displays the correlation between low birth-weight on the y-axis and ozone on the x-axis. In the exploratory models, I found the relationship between ozone and low birth-weight to be the strongest, so it is useful to display the distribution of these two variables together. Though ozone is a continuous variable, the data has been recorded in such a way that it only takes certain values rounded to the hundredth unit—-this is why the census tracts are plotted in a pattern of vertical lines. Because there is such a large number of census tract observations being plotted, it is not immediately apparent that there is any relationship between these two variables. However, when a trend line is added to the plot, a slight positive relationship can be perceived in the data. Census tracts with higher recorded ozone levels also have higher rates of low birth-weight, on average.

``` {r descriptive statistics—-LBW distribution, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}

# Univariate distribution of LBW
hist(ces3_complete$lbw, main="Fig 1. Low Birth-Weight by Census Tract", xlab="% of births", ylab="# of Census Tracts", col="hotpink")
summary(ces3_complete$lbw)

```

``` {r descriptive statistics--ozone and LBW, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
# Joint distribution of ozone and LBW
plot(ces3_complete$ozone, ces3_complete$lbw, main="Fig 2. Census Tracts by Ozone and Low Birth-Weight", xlab="Ozone", ylab ="Low Birth-Weight %")
abline(lm(ces3_complete$lbw~ces3_complete$ozone))

```

# Results

Four regression models were constructed for this analysis. Note that because the ozone data only ranges from 0.02 to 0.068 parts per million, it is more meaningful to interpret the coefficients for this variable by considering the effect of a 1/100th-unit increase, rather than a 1-unit increase. Therefore, in the following interrpretations, I divide the ozone coefficient by 100.

Model 1 is a "kitchen sink" model which includes all of the environmental effects variables added together. The point of this model was to determine which pollution variables are associated most stronlgy with LBW. Ozone, PM2.5, and diesel PM are statistically significant at the p<0.01 level. A .01 ppm increase of ozone is associated with a .17 percentage point increase of LBW in the tracts. This is the strongest predictor of LBW. A 1-μg/m3 increase in PM2.5 is associated with a 0.05 percentage point increase in LBW, and a 1-kg/day increase in diesel PM is associated with a 0.01 increase in LBW. Several other pollutants (traffic, cleanup sites, impaired water bodies, and solid waste) are statistically significant at the .05 level, but the magnitude of the coefficients is very small. These variables do not seem to be playing a major role in low birth-weight outcomes. This, in combination with the results of models 3 and 4, is my rationale for focusing on ozone, PM2.5, and diesel PM in my analysis.

Model 2 considers the independent effect of poverty on LBW. The effect is statistically significant at the .001 level, and its magnitude is fairly large: for every percentage point increase in the poverty level, the model predicts a .02 percentage point increase in LBW, on average. Inclusion of poverty in the model reduces the magnitude of the key pollution variables identified in model 1. The predictive strength of ozone drops from .17 to .06, pm2.5 from .05 to .04, and diesel PM from .01 to .008. All three variables remain statistically significant at the same levels, however. This means that while poverty exerts an independent effect on the incidence of LBW, and is substantially collinear with the pollution variables, it does not completely account for the association of environmental pollution with LBW in census tracts.

Model 3 assesses the independent effect of race on LBW. This model indicates some surprising findings. It predicts that for every 1 percentage point increase in the nonwhite population in census tracts, there will be a .02 percentage point increase in LBW. The race variable in this model is statistically significant at the .001 level. The inclusion of race changes some predicted associations of environmental pollutants and LBW present in the earlier models. The effect of ozone actually increases with the inclusion of race, from .17 units in model 1 to .20 here [ask Aaron about this]. The effect of diesel PM pollution remains 0.01. PM 2.5 pollution, however, is no longer statistically significant in this model, and the magnitude of its effect has dropped to almost 0, from .05 in model 1 to -0.01. PM 2.5's statistical insignificance in this model suggests that a higher nonwhite population in census tracts accounts for the association of PM2.5 pollution with LBW in models 1 and 2. Similar to model 1, ozone and diesel PM remain stasticially significant. Some of the other environmental pollutants——drinking water, pesticides, and toxic releases——are statistically significant in this model, but their magnitudes remain small.

Model 4 includes both the effects of poverty and the nonwhite population. Both effects are statistically significant. The indepdendent effect of poverty here has dropped from .02 to .01; the effect of percent-nonwhite population has stayed the same, as .02. PM2.5 pollution seems to be very highly correlated with race. Because PM2.5 went from statistically significant in model 2 to not statistically significant in the models including a race variable, it seems that this pollution variable is fully explained by race, whereas the other pollution variables——which retain statistical significance and a significant magnitude in these models——are not. Overall, the results of this model suggest that race and poverty each appear to have an effect on LBW outcomes, but this effect does not fully predict the level of LBW. Environmental effects, especially ozone pollution, retain a significant independent influence on the rate of LBW, even when the confounding roles of poverty and race are accounted for.

``` {r Models, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}

model1 <- lm(lbw~ozone+pm2.5+dieselpm+drinkingwater+pesticides+toxrelease+traffic+
               cleanupsite+groundwater+hazwastes+impwater+solidwaste, data=ces3_complete)

model2 <- lm(lbw~ozone+pm2.5+dieselpm+drinkingwater+pesticides+toxrelease+traffic+
               cleanupsite+groundwater+hazwastes+impwater+solidwaste+poverty, data=ces3_complete)

model3 <- lm(lbw~ozone+pm2.5+dieselpm+drinkingwater+pesticides+toxrelease+traffic+
               cleanupsite+groundwater+hazwastes+impwater+solidwaste+pctnonwhite, data=ces3_complete)

model4 <- lm(lbw~ozone+pm2.5+dieselpm+drinkingwater+pesticides+toxrelease+traffic+
               cleanupsite+groundwater+hazwastes+impwater+solidwaste+poverty+pctnonwhite, data=ces3_complete)

library(stargazer)
stargazer(model1, model2, model3, model4, type="text", digits = 3, covariate.labels = c("Ozone", "PM2.5", "Diesel PM", "Drinking Water Quality", "Pesticides", "Toxic Releases", "Traffic", "Proximity to Cleanup Sites", "Groundwater", "Hazardous Wastes", "Impaired Water", "Solid Waste", "Poverty", "Nonwhite Tract Percentage"), dep.var.labels = "Low Birth-Weight")

```

# Discussion

In this analysis, I have applied an OLS regression method to examine the influence of various environmental and social stressors upon low birth-weight (LBW) outcomes in the CalEnviroScreen 3.0 data analysis tool. Four models were constructed to assess the relative strengh of correlation between low birth-weight and variables like ozone, particulate matter air pollution, hazardous wastes, and social conditions like poverty or race. From this analysis, three findings seem particularly important. First, air pollution, including ozone, PM2.5 pollution, and diesel PM, has a stronger effect on LBW outcomes in California census tracts than other forms of pollution, such as proximity to hazardous waste sites, impaired water quality, or pesticide exposure. While this analysis doesn't allow me to determine the precise causal mechanism behind the different strength of association for these variables, it seems reasonable to assume that air pollution has a stronger effect on many kind of health outcomes, including LBW, because of its very wide geographical dispersion. Other forms of pollution, such as exposure to pesticides or hazardous wastes, have a strong demonstrated effect upon adverse birth outcomse at the local level, but do not pose nearly as great a risk over a wide geographic area.

The models which include poverty and race controls reveal two other interesting findings. First, when race is controlled for (in model 3), PM2.5 is no longer statistically significant, and the magnitude of its effect drops significantly. This indicates that a higher nonwhite population in census tracts accounts for the association of PM2.5 pollution with LBW present in the model which does not include any race variable. Second, the model which controls for both poverty and race indicates that while the inclusion of these variables reduces the magnitude and statistical significance of PM2.5, the other air pollution effects, ozone and diesel PM, retain a significant independent effect on LBW. This is especially true of ozone. The role of ozone is particularly interesting in this analysis. In model 2, when poverty is controlled for, the effect of ozone diminishes considerably. In model 3, however, when race is controlled for, the effect of ozone actually increases from its level in model 1. Finally, in model 4, its effect returns almost exactly to its level in model 1. It is not clear to me why this is the case. Future research could explore the particular relationship of ozone in the presence or absence of different, or more disaggregated, socio-economic variables of this kind.

The main implication of this study seems to be that, among all major environmental stressors, the most significant threats to public health are air pollution, especially ozone. Policies aimed at reducing adverse birth outcomes, which have a disproportionate effect upon non-white populations, should prioritize efforts to reduce these pervasive forms of pollution in California and elsewhere. Doing so would have a particularly positive effect for poor and non-white people. However, this analysis also indicates that there is an independent effect of poverty and racialization above and beyond the effects of air pollution. Policies aimed at reducing poor birth outcomes will have to take this into account.

This analysis has a number of important limitations. First, this data only includes census tracts in California. While results in a state as large and socially diverse as California are plausibly generalizable to many other contexts, future research on pollution and LBW should explore the relationship in other contexts and at different scales. Second, future analyses could use more complex methods to explore interactions between different variables. Third, more detailed socio-economic data than is available in CalEnviroScreen could help to explain puzzles revealed in this analysis, such as the up-and-down strength of the ozone effect on LBW.

# References

Barker, D. J. P., J. G. Eriksson, T. Forsén, and C. Osmond. 2002. “Fetal Origins of Adult Disease: Strength of Effects and Biological Basis.”       International Journal of Epidemiology 31(6):1235–39. doi: 10.1093/ije/31.6.1235.
Bekkar, Bruce, Susan Pacheco, Rupa Basu, and Nathaniel DeNicola. 2020. “Association of Air Pollution and Heat Exposure With Preterm Birth, Low Birth Weight, and Stillbirth in the US: A Systematic Review.” JAMA Network Open 3(6):e208243. doi: 10.1001/jamanetworkopen.2020.8243.
Berry, M., and F. Bove. 1997. “Birth Weight Reduction Associated with Residence near a Hazardous Waste Landfill.” Environmental Health Perspectives 105(8):856–61.
Bhutta, Adnan T., Mario A. Cleves, Patrick H. Casey, Mary M. Cradock, and K. J. S. Anand. 2002. “Cognitive and Behavioral Outcomes of School-Aged Children Who Were Born Preterm: A Meta-Analysis.” JAMA 288(6):728–37. doi: 10.1001/jama.288.6.728.
Currie, Janet, Joshua Graff Zivin, Katherine Meckel, Matthew Neidell, and Wolfram Schlenker. 2013. “Something in the Water: Contaminated Drinking Water and Infant Health.” The Canadian Journal of Economics. Revue Canadienne d’economique 46(3):791–810. doi: 10.1111/caje.12039.
Ghosh, Jo Kay C., Michelle Wilhelm, Jason Su, Daniel Goldberg, Myles Cockburn, Michael Jerrett, and Beate Ritz. 2012. “Assessing the Influence of Traffic-Related Air Pollution on Risk of Term Low Birth Weight on the Basis of Land-Use-Based Regression Models and Measures of Air Toxics.” American Journal of Epidemiology 175(12):1262–74. doi: 10.1093/aje/kwr469.
Greenfield, Ben K., Jayant Rajan, and Thomas E. McKone. 2017. “A Multivariate Analysis of CalEnviroScreen: Comparing Environmental and Socioeconomic Stressors versus Chronic Disease.” Environmental Health 16. doi: 10.1186/s12940-017-0344-z.
Harley, Kim G., Karen Huen, Raul Aguilar Schall, Nina T. Holland, Asa Bradman, Dana Boyd Barr, and Brenda Eskenazi. 2011. “Association of Organophosphate Pesticide Exposure and Paraoxonase with Birth Outcome in Mexican-American Women.” PLOS ONE 6(8):e23923. doi: 10.1371/journal.pone.0023923.
Laurent, Olivier, Jun Wu, Lianfa Li, Judith Chung, and Scott Bartell. 2013. “Investigating the Association between Birth Weight and Complementary Air Pollution Metrics: A Cohort Study.” Environmental Health 12(1):18. doi: 10.1186/1476-069X-12-18.
Laurent, Olivier, Jianlin Hu, Lianfa Li, Michael J. Kleeman, Scott M. Bartell, Myles Cockburn, Loraine Escobedo, and Jun Wu. 2016. “A Statewide Nested Case–Control Study of Preterm Birth and Air Pollution by Source and Composition: California, 2001–2008.” Environmental Health Perspectives 124(9):1479–86. doi: 10.1289/ehp.1510133.
Lu, Michael C., and Neal Halfon. 2003. “Racial and Ethnic Disparities in Birth Outcomes: A Life-Course Perspective.” Maternal and Child Health Journal 7(1):13–30. doi: 10.1023/A:1022537516969.
Martinson, Melissa L., and Nancy E. Reichman. 2016. “Socioeconomic Inequalities in Low Birth Weight in the United States, the United Kingdom, Canada, and Australia.” American Journal of Public Health 106(4):748–54. doi: 10.2105/AJPH.2015.303007.
Nepomnyaschy, Lenna, and Nancy E. Reichman. 2006. “Low Birthweight and Asthma Among Young Urban Children.” American Journal of Public Health 96(9):1604–10. doi: 10.2105/AJPH.2005.079400.
Ponce, Ninez A., Katherine J. Hoggatt, Michelle Wilhelm, and Beate Ritz. 2005. “Preterm Birth: The Interaction of Traffic-Related Air Pollution with Economic Hardship in Los Angeles Neighborhoods.” American Journal of Epidemiology 162(2):140–48. doi: 10.1093/aje/kwi173.
Rodriquez, Matthew, and Lauren Zeise. 2017. CalEnviroScreen 3.0 Sourcebook. Sacramento, CA: California Office of Environmental Health Hazard Assessment.
Sims, Mario, Tammy L. Sims, and Marino A. Bruce. 2008. “Race, Ethnicity, Concentrated Poverty, and Low Birth Weight Disparities.” Journal of National Black Nurses’ Association : JNBNA 19(1):12–18.
World Health Organization (WHO). 2018. “Preterm Birth.” Retrieved October 19, 2020 (https://www.who.int/news-room/fact-sheets/detail/preterm-birth).